The Two Dualities of Computation: Negative and Fractional Types

### 2012-03-11 rational.tex
sat

The construction presented here is a novel SAT-solver that relies on
{{trace*}} for its solution. It varies from previous classical SAT solvers
\amr{must cite ???} in that there is no explicit search operation on the
boolean space. It also varies from previous quantum SAT solvers \amr{must
  cite ???} in that it does uses a {{trace}} operation to compute a fixpoint
which is the solution following which an isomorphic clone operation allows us
to examine the result.  Though we focus on SAT here, this construction can be
generalized to any constraint satisfaction problem whose search space can be
represented by a recursive type and for which an isomorphic clone operation
can be constructed. Before we start, let us recall a constructions from 
prior work \cite{infeffects}.

\emph{Conditionals.}
Given any combinator {{c : b <-> b}} we can construct a combinator called
{{if_c : bool*b <->bool*b}} in terms of {{c}}, where {{if_c}} behaves like a
one-armed $\mathit{if}$-expression. If the supplied boolean is {{true}} then
the combinator {{c}} is used to transform the value of type~{{b}}. If the
boolean is {{false}}, then the value of type {{b}} remains unchanged. We can
write down the combinator for {{if_c}} in terms of {{c}} as 
{{ dist (;) ((id (*) c) (+) id) (;) factor }}.

@@

%% \begin{enumerate}
%% \item Recall from previous work that for the types {{bool}} and, by
%%   extension, any tuple of {{bool}}, an isomorphic clone operation can
%%   be constructed (See Lemma 7.2 \cite{infeffects}).

%% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% \item Also, Sec. 3.3 \cite{infeffects} constructed the combinator
%%   {{if_{not} }}, which when generalized to two bits was
%%   {{toffoli}}. This construction maybe generalised to {{n}} bits as
%%   {{if^n_{not} }}.
%% \end{enumerate}

%% That said, we will focus on the limited case of boolean satisfiability
%% here.

\emph{A Total Annihilation Circuit.}  Consider the circuit below that
has no observable output, i.e., for all inputs it is
annihilating. This happens because it constructs a boolean and its
dual, negates one of them and attempts to collapse them.

>>>>

The construction presented here is a novel SAT-solver that relies on
{{trace*}} for its solution. It varies from previous classical SAT solvers
\amr{must cite ???} in that there is no explicit search operation on the
boolean space. It also varies from previous quantum SAT solvers \amr{must
  cite ???} in that it does uses a {{trace}} operation to compute a fixpoint
which is the solution following which an isomorphic clone operation allows us
to examine the result.  Though we focus on SAT here, this construction can be
generalized to any constraint satisfaction problem whose search space can be
represented by a recursive type and for which an isomorphic clone operation
can be constructed.

\subsection{Constructions with Booleans}

Before we start, let us recall a construction of conditionals from
prior work \cite{infeffects}.  Given any combinator {{c : b <-> b}} we
can construct a combinator called {{if_c : bool*b <->bool*b}} in terms
of {{c}}, where {{if_c}} behaves like a one-armed
$\mathit{if}$-expression. If the supplied boolean is {{true}} then the
combinator {{c}} is used to transform the value of type~{{b}}. If the
boolean is {{false}}, then the value of type {{b}} remains
unchanged. We can write down the combinator for {{if_c}} in terms of
{{c}} as {{ dist (;) ((id (*) c) (+) id) (;) factor }}.

@@

DIAGRAM clone

DIAGRAM clone n bits

\subsection{Construction of the Solver}
The key insight underlying the construction comes from the fact that we
can build \emph{annihilation circuits}.  Consider the circuit below
that has no observable output, i.e., for all inputs it is
annihilating. This happens because it constructs a boolean and its
dual, negates one of them and attempts to collapse them.
